### 2017-08-13 DCMIP2016-Part1-EditorResponse.tex

\documentclass[11pt]{article}

\begin{document}

\noindent \textbf{DCMIP2016:  A Review of Non-hydrostatic Dynamical Core Design and Intercomparison of Participating Models} \\

\noindent {Paul A. Ullrich, Christiane Jablonowski, James Kent, Peter H. Lauritzen, Ramachandran Nair, Kevin A. Reed, Colin M. Zarzycki, David M. Hall, Don Dazlich, Ross Heikes, Celal Konor, David Randall, Thomas Dubos, Yann Meurdesoif, Xi Chen, Lucas Harris, Christian K\"uhnlein, Vivian Lee, Abdessamad Qaddouri, Claude Girard, Marco Giorgetta, Daniel Reinert, Joseph Klemp, Sang-Hun Park, William Skamarock, Hiroaki Miura, Tomoki Ohno,
Ryuji Yoshida, Robert Walko, Alex Reinecke, and Kevin Viner} \\

\noindent \textbf{Response to Editor} \\

\noindent A throughly revised code availability table has been included with the latest version of the paper.  Six of the modeling groups have uploaded model source code to Zenodo.  Licensing agreements prevent the remainder from directly posting model code publicly, however nearly all provide an avenue for public access to the codebase.  Only two of the model groups (ACME-A and FVM) cannot allow public access at this time, but plan for a public release of the code in the future.  In the mean time, these groups will allow the topical editor (Julia Hargreaves) access the code.  The following page contains the updated code availability table.

\end{document}